#Chapter 4

*Clustering and classification*

First I loaded the Boston data from MASS package.

```{r}
library(MASS)
data("Boston")
dim(Boston)
```
You can see that the Boston-data is a dataset in which there are 506 observations of 14 variables. The Boston-data is about housing values in suburbs of Boston.
```{r}
str(Boston)
```
The variables vary from crime rate per capita to whether the suburb bounds river Charles. 
```{r}
summary(Boston)
```
Summary of the data. 

```{r}
boston_scaled <- scale(Boston)
boston_scaled <- as.data.frame(boston_scaled)
summary(boston_scaled)

```
Using the scale() -function I standardized the data. Now every single variable has mean of 0.

```{r}
scaled_crim <- boston_scaled$crim
bins <- quantile(scaled_crim)
crime <- cut(scaled_crim, breaks = bins, include.lowest = TRUE)
boston_scaled <- dplyr::select(boston_scaled, -crim)
boston_scaled <- data.frame(boston_scaled, crime)
n <- nrow(boston_scaled)
ind <- sample(n,  size = n * 0.8)
train <- boston_scaled[ind,]
test <- boston_scaled[-ind,]

```
Code above creates the train and test datas.

```{r}
lda.fit <- lda(crime ~ . , data = train)
lda.fit
classes <- as.numeric(train$crime)
plot(lda.fit, dimen = 2,col = classes, pch = classes)
```
This is the linear discriminant analysis biplot.

```{r}
correct_classes <- test$crime
test <- dplyr::select(test, -crime)
lda.pred <- predict(lda.fit, newdata = test)
table(correct = correct_classes, predicted = lda.pred$class)
```

