#Chapter 5

*Dimensionality reduction techniques*

First of all ignore all these warning messages. I couldn't get rid of those.

```{r}
setwd("~/IODS/IODS-project/data")

human <- read.csv("human.csv", stringsAsFactors = F, sep=",", row.names = 1)

library(ggplot2)
library(dplyr)
library(stringr)
library(GGally)

```
First I loaded the data and necessary packages.  

```{r}
str(human)
dim(human)
```
Above you can see the structure and dimensions of the human.csv data. This data has 155 observations of 8 variables. Its variables describe human values such as life expectancy and expected length of education.  

```{r}
ggpairs(human)
```
Above you can see visualization of the data. The variables have surprisingly large correlation values between variables. For example -0.857 correlation between maternal mortality and life expectancy. Expected lenght of education seems to be quite normally distributed.  

```{r}
pca_human <- prcomp(human)
summary(pca_human)

```

Above you can see the summary of the principal components.

```{r}
biplot(pca_human, choices = 1:2, cex = c(0.8, 1), col = c("grey40", "deeppink2"))
```

You really can't tell much of the non-standardized biplot. Its a mess. Now let's standardize the data and plot it.

```{r}
human_std <- scale(human)
pca_human <- prcomp(human_std)
biplot(pca_human, choices = 1:2, cex = c(0.8, 1), col = c("grey40", "deeppink2"))

```

As you can see the distribution is now distributed more all-around. This is because the data has been changed to a standard normal distribution. 
